---
title: '220322'
author: "Karim Primov"
date: '2022-03-22'
output: html_document
---

```{r}
library(tidyverse)
library(manipulate)
library(ggplot2)
library(patchwork)
library(broom)
library(infer)
```

```{r}
f <-  "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/zombies.csv"
d <- read_csv(f, col_names = TRUE)
```

```{r}
d <- mutate(d, centered_height = height - mean(height))
d <- mutate(d, centered_weight = weight - mean(weight))

p1 <- ggplot(data = d, aes(x = weight, y = height)) + geom_point()
p2 <- ggplot(data = d, aes(x = centered_weight,
  y = centered_height)) + geom_point()

p1 + p2
```

```{r}
slope.test <- function(beta1, data){
  g <- ggplot(data=data, aes(x = centered_weight, y = centered_height))
  g <- g + geom_point()
  g <- g + geom_abline(intercept = 0, slope = beta1, size = 1, colour="blue", alpha=1/2)
  ols <- sum((data$centered_height - beta1 * data$centered_weight) ^2)
  g <- g + ggtitle(paste("Slope = ", beta1,
    "\nSum of Squared Deviations = ", round(ols, 3)))
  g
}
```

```{r}
manipulate(slope.test(beta1, data=d),
  beta1 = slider(-1, 1, initial = 0, step = 0.005))
```

```{r}
b1 <- cor(d$height, d$weight) * sd(d$height)/sd(d$weight)
b1
```

```{r}
b0 <- mean(d$height) - b1*mean(d$weight)
b0
```

```{r}
model <- lm(formula = height~ weight, data = d)
model
summary(model)
```

```{r}
f <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/Street_et_al_2017.csv"
d <- read_csv(f, col_names = TRUE)
```

```{r}
library(skimr)
skim(d)
```
Plot brain size as a function of social group size, longevity, juvenile period length, and reproductive life span 

Brain size~social group size 
```{r}
d <- mutate(d, centered_ECV = ECV - mean(ECV))
d <- mutate(d, centered_Group_size = Group_size - mean(Group_size))

p1 <- ggplot(data = d, aes(x = Group_size, y = ECV)) + geom_point()


p1
```

Brain size~longevity
```{r}
d <- mutate(d, centered_ECV = ECV - mean(ECV))
d <- mutate(d, centered_Longevity = Longevity - mean(Longevity))

p1 <- ggplot(data = d, aes(x = Longevity, y = ECV)) + geom_point()


p1 
```

Brain size~juvenile period length(which variable is juvenile period length?)
```{r}
d <- mutate(d, centered_ECV = ECV - mean(ECV))
d <- mutate(d, centered_Group_size = Group_size - mean(Group_size))

p1 <- ggplot(data = d, aes(x = Group_size, y = ECV)) + geom_point()


p1
```

Brain size~reproductive life span
```{r}
d <- mutate(d, centered_ECV = ECV - mean(ECV))
d <- mutate(d, centered_Repro_lifespan = Repro_lifespan - mean(Repro_lifespan))

p1 <- ggplot(data = d, aes(x = Repro_lifespan, y = ECV)) + geom_point()


p1 
```

Calculate ordinary least squares regression coefficients B1 and B0 for ECV ~ social group size

Removing NAs
```{r}
d_mod <- d %>% filter(!is.na(ECV) & !is.na(Group_size))
```

```{r}
b1.nonas <- cor(d_mod$ECV, d_mod$Group_size) * sd(d_mod$ECV)/sd(d_mod$Group_size)
b1.nonas
```

```{r}
b0.nonas <- mean(d_mod$ECV) - b1.nonas*mean(d_mod$Group_size)
b0.nonas
```

Calculating regression coefficients using 'lm' function
```{r}
model <- lm(formula = ECV~ Group_size, data = d_mod)
model
summary(model)
```



For catarrhines:

Calculate ordinary least squares regression coefficients B1 and B0 for ECV ~ social group size

Removing NAs
```{r}
d_mod_cat <- d %>% filter(!is.na(ECV) & !is.na(Group_size) & Taxonomic_group=="Catarrhini")
```

```{r}
b1 <- cor(d_mod_cat$ECV, d_mod_cat$Group_size) * sd(d_mod_cat$ECV)/sd(d_mod_cat$Group_size)
b1
```

```{r}
b0 <- mean(d_mod_cat$ECV) - b1*mean(d_mod_cat$Group_size)
b0
```

Calculating regression coefficients using 'lm' function
```{r}
model <- lm(formula = ECV~ Group_size, data = d_mod_cat)
model
summary(model)
```

For platyrrhines:

Calculate ordinary least squares regression coefficients B1 and B0 for ECV ~ social group size

Removing NAs
```{r}
d_mod_plat <- d %>% filter(!is.na(ECV) & !is.na(Group_size) & Taxonomic_group=="Platyrrhini")
```

```{r}
b1 <- cor(d_mod_plat$ECV, d_mod_plat$Group_size) * sd(d_mod_plat$ECV)/sd(d_mod_plat$Group_size)
b1
```

```{r}
b0 <- mean(d_mod_plat$ECV) - b1*mean(d_mod_plat$Group_size)
b0
```

Calculating regression coefficients using 'lm' function
```{r}
model <- lm(formula = ECV~ Group_size, data = d_mod_plat)
model
summary(model)
```

For strepsirhines:

Calculate ordinary least squares regression coefficients B1 and B0 for ECV ~ social group size

Removing NAs
```{r}
d_mod_strep <- d %>% filter(!is.na(ECV) & !is.na(Group_size) & Taxonomic_group=="Strepsirhini")
```

```{r}
b1 <- cor(d_mod_strep$ECV, d_mod_strep$Group_size) * sd(d_mod_strep$ECV)/sd(d_mod_strep$Group_size)
b1
```

```{r}
b0 <- mean(d_mod_strep$ECV) - b1*mean(d_mod_strep$Group_size)
b0
```

Calculating regression coefficients using 'lm' function
```{r}
model <- lm(formula = ECV~ Group_size, data = d_mod_strep)
model
summary(model)
```


Calculate standard error for the slope coefficient, the 95% CI, and p-value associated with this coefficient by hand
```{r}
m <- lm(data = d_mod, ECV ~ Group_size)
summary(m)
```

CI using confint function
```{r}
CI <- confint(m, level = 1-0.05)
CI
```

Getting m.summary
```{r}
m.summary <- tidy(m)
m.summary
```

Calculate standard error by hand
```{r}
n <- nrow(d_mod)
mean_x <- mean(d_mod$Group_size)
y_pred <- b0.nonas + b1.nonas*d_mod$Group_size
y_error = d_mod$ECV - y_pred

num <- sum(y_error^2)
den <- (n-2)* sum((d_mod$Group_size - mean_x)^2)

se <- sqrt(num/den)
se

```


Calculate 95% CI by hand
```{r}
# by hand
lower <- m.summary$estimate -
  qt(1 - alpha / 2, df = nrow(d) - 2) * m.summary$std.error
upper <- m.summary$estimate +
  qt(1 - alpha / 2, df = nrow(d) - 2) * m.summary$std.error
CI <- cbind(lower, upper)
rownames(CI) <- c("(Intercept)", "Group_size")
colnames(CI) <- c(paste0(as.character(alpha/2 * 100), " %"),paste0(as.character((1-alpha/2) * 100), " %"))
CI
```

Calculate p-value by hand
```{r}
m.summary$calc.statistic <- (m.summary$estimate-0)/m.summary$std.error 
m.summary$calc.p.value <- 2 * pt(m.summary$calc.statistic,
  df=nrow(d)-2, lower.tail = FALSE)
m.summary
```


Use a permutation approach with 1000 permutations to generate a null sampling distribution for slope coefficient. What is it that you need to permute? What is the p-value associated with your original slope coefficient?
```{r}
permuted.slope <- d_mod %>%
  # specify model
  specify(ECV ~ Group_size) %>%
  # use a null hypothesis of independence
  hypothesize(null = "independence") %>%
  # generate permutation replicates
  generate(reps = 1000, type = "permute") %>%
  # calculate the slope statistic
  calculate(stat = "slope")

head(permuted.slope)
```
To generate the null sampling distribution, you need to permute values of the response variable (e.g. ECV) under null hypothesis that the response variable is independent of the explanatory variable (e.g. group size). 

Calculate original slope, then obtain associated p-value
```{r}
#Original Slope Calculation
# first define alpha, CI boundaries, and critical values
alpha <- 0.05
confidence_level <- 1 - alpha
p_lower <- alpha / 2
p_upper <- 1 - (alpha / 2)
degrees_of_freedom <- nrow(d) - 2
critical_value <- qt(p_upper, df = degrees_of_freedom)

# original slope
original.slope <- lm(data = d_mod, ECV ~ Group_size) %>%
  # tidy the model and add the CI based on the t distribution
  tidy(conf.int=TRUE, conf.level=confidence_level) %>%
  # or manually calculate the CI based on the t distribution
  mutate(
    lower = estimate - std.error * critical_value,
    upper = estimate + std.error * critical_value
  ) %>%
  filter(term=="Group_size")
original.slope # show model results for slope of Group_size
```

```{r}
#p-value calculation
p.value <- permuted.slope %>% 
  # add a column of the absolute value of the slope
  mutate(abs_stat=abs(stat)) %>%
  # calculate a summary statistic
  summarize(
    # calculate proportion of cases where the absolute value
    # of the permuted slope is greater than or equal to the 
    # absolute value of the observed slope
    estimate = mean(abs_stat >= abs(pull(original.slope, estimate)))
  )

p.value
```



Use bootstrapping to generate 95% CI for your estimate of slope coefficient using both percentile method and theory-based method.


Generating bootstraps
```{r}
boot.slope <- d_mod %>%
  # specify model
  specify(ECV ~ Group_size) %>%
  # generate bootstrap replicates
  generate(reps = 1000, type = "bootstrap") %>%
  # calculate the slope statistic
  calculate(stat = "slope")

head(boot.slope)
```

Creating CIs for regression coefficients using two methods
```{r}
boot.slope.summary <- boot.slope %>%
  # summarize the mean, t distribution based CI, and quantile-based CI
  summarize(
    # mean of stat
    estimate = mean(stat),
    # std error of stat
    std.error = sd(stat),
    # calculate the CI based on the SE and t distribution
    boot.se.method.low = estimate - std.error * critical_value,
    boot.se.method.high = estimate + std.error * critical_value,
    # calculate the CI based on the quantile (percentile)  method
    boot.percent.method.low = quantile(stat, p_lower),
    boot.percent.method.high = quantile(stat, p_upper)
  )
# show summary of bootstrap sampling distribution
boot.slope.summary
```


What is the p-value associated with your observed slope coefficient based on each of these methods? 


P-values based on percentile method
```{r}
# first define alpha, CI boundaries, and critical values
alpha <- 0.05
confidence_level <- 1 - alpha
p_lower <- alpha / 2
p_upper <- 1 - (alpha / 2)
degrees_of_freedom <- nrow(d) - 2
critical_value <- qt(p_upper, df = degrees_of_freedom)

# original slope
original.slope <- lm(data = d_mod, ECV ~ Group_size) %>%
  # tidy the model and add the CI based on the t distribution
  tidy(conf.int=TRUE, conf.level=confidence_level) %>%
  # or manually calculate the CI based on the t distribution
  mutate(
    lower = estimate - std.error * critical_value,
    upper = estimate + std.error * critical_value
  ) %>%
  filter(term=="Group_size")
original.slope # show model results for slope of Group_size
```


P-values based on theoretical method
```{r}
# first define alpha, CI boundaries, and critical values
alpha <- 0.05
confidence_level <- 1 - alpha
p_lower <- alpha / 2
p_upper <- 1 - (alpha / 2)
degrees_of_freedom <- nrow(d) - 2
critical_value <- qt(p_upper, df = degrees_of_freedom)

# original slope
original.slope <- lm(data = d_mod, ECV ~ Group_size) %>%
  # tidy the model and add the CI based on the t distribution
  tidy(conf.int=TRUE, conf.level=confidence_level) %>%
  # or manually calculate the CI based on the t distribution
  mutate(
    lower = quantile(stat, p_lower),
    upper = quantile(stat, p_upper)
  ) %>%
  filter(term=="Group_size")
original.slope # show model results for slope of Group_size
```








```{r}

```

```{r}

```

```{r}

```

```{r}

```